Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-19 AE_Reponse.tex
FINAL PASS THROUGH ALL DOCUMENTS

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
%When an algorithm has warm starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
%Depending on the effectiveness of warm starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
%\medskip 
In addition, based on the comments of Referee 1, we studied efficiency of overlapping estimators compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with equivalently lower variances. 
Our results indicate that when warm starting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warm starting is quite low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\bigskip

\noi 
We illustrated these results on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping levels, and (iii) this analysis provides a way to select the degree of overlapping by maximizing efficiency.\bigskip

\noi
Finally, we provided guidelines on selecting the degree of overlap at the end of \S 5.4 (Computational Efficiency) based on our study. 
We believe the paper has improved as a result of the study of computational efficiency. 
Thank you for this suggestion. 
\medskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem using insertion sort.
We also (i) changed \S 1 (Introduction) to clearly motivate the study and argue usefulness of the results, paying specific attention to efficiency gains; (ii) updated \S 6 (Conclusions) to summarize these new results, again paying attention to computational efficiency; (iii) provided a study of computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming in the new \S 5.4 (Computational Efficiency); and finally (iv) updated the abstract.  
%(Please see our response to the previous comment.)
\medskip

@@

\noi
This sentence was removed as a result of the computational efficiency study. 
Based on the examination of computational efficiency, we provided new guidelines on selecting the degree of overlapping.

>>>>

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
%When an algorithm has warm starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We analyzed efficiency for a given functional form of the solution time per batch. 
%Depending on the effectiveness of warm starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
%\medskip 
In addition, based on the comments of Referee 1, we studied the efficiency of OMRP compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with equivalently lower variances. 
Our results indicate that when warm starting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warm starting is low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\medskip

\noi 
We illustrated these results on the newsvendor problem using insertion sort. 
Our experiments show that (i) the functional form used to express solution time per batch provides a good representation, (ii) increased computational efficiency over the nonoverlapping counterpart is indeed observed over a range of overlapping degrees, and (iii) this analysis provides a way to select the degree of overlap by maximizing efficiency and then minimizing overall computation time.\bigskip

\noi
As a result of this study, we provided guidelines on selecting the degree of overlap in OMRP at the end of \S 5.4 (Computational Efficiency).  
We believe the paper has substantially improved as a result of the study of computational efficiency. 
Thank you for this suggestion. 
\medskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem using insertion sort.
We also studied the computational efficiency of overlapping for assessment of solution quality in stochastic programming in the new \S 5.4 (Computational Efficiency), based on your earlier comments. 
In addition, we (i) changed \S 1 (Introduction) to clearly motivate the study and argue usefulness of the results, (ii) updated \S 6 (Conclusions) to summarize these new results, paying specific attention to computational efficiency, and finally (iv) updated the abstract.  
%(Please see our response to the previous comment.)
\medskip

@@

\noi
This sentence was removed as a result of the computational efficiency study. 
Based on the examination of computational efficiency, we provided new guidelines on selecting the degree of overlap.

### 2014-07-19 EinC_Reponse.tex
FINAL PASS THROUGH ALL DOCUMENTS

\noindent  
We incorporated all the recommended changes suggested by you, the associate editor, and the referees and believe the paper is significantly improved as a result.
We very much appreciate everyone's time and constructive comments. 
As a result of the revision, the wording and content of the paper, compared to the Winter Simulation Conference Proceedings version, has partly changed.
We also edited the paper to improve exposition and tried to cut back word for word copying during this process.  
\medskip

@@

\noindent
We added a paragraph in Section 1 (Introduction) detailing what has been done over the WSC proceedings paper. 
Please see the first paragraph on page 5 of the revised manuscript.\bigskip

>>>>

\noindent  
We incorporated all the recommended changes suggested by you, the associate editor, and the referees and believe the paper is significantly improved as a result.
We very much appreciate everyone's time and constructive comments. 
As a result of the revision, the wording and content of the paper, compared to the Winter Simulation Conference Proceedings version, has changed.
We also edited the paper to improve exposition and tried to cut back word for word copying during this process.  
\medskip

@@

\noindent
We added a paragraph in \S 1 (Introduction) detailing what has been done over the WSC proceedings paper. 
Please see the first paragraph on page five of the revised manuscript.\bigskip

### 2014-07-19 omrp_lb_rev.tex
FINAL PASS THROUGH ALL DOCUMENTS

\begin{figure}[thb!]
	\centering
	\subfloat[][Theoretical]{
	   \label{fig:efficiency_theory}
	   \includegraphics[width=.49\textwidth]{images/comp_efficiency_theory}}
	\subfloat[][Empirical for NVD]{
	   \label{fig:efficiency_nv}
	   \includegraphics[width=.51\textwidth]{images/nv_comp_efficiency}}
		\caption{
		Computational efficiency of the OMRP variance estimator relative to the (nonoverlaping) MRP variance estimator measured
		(a) theoretically and
		(b) empirically for NVD using insertion sort
%		(graphs are scaled so that nonoverlapping batches produces a computational efficiency of 1)
		}
\label{fig:efficiency}
\end{figure}

@@

We have applied the overlapping batch means method from simulation output analysis \citep{Meketon1984,Song1992,Welch1987} to assessing solution quality in stochastic programming, in particular, to Monte Carlo simulation-based estimators of optimality gaps \citep{Mak1999}. 
We have provided conditions under which (i) the resulting point estimator of the optimality gap and its associated variance estimator are consistent, (ii) OMRP variance estimators achieve asymptotically lower variances relative to the MRP variance estimator, and (iii) OMRP CIs are asymptotically valid. 
Empirical results indicate that asymptotic reductions in variance of the variance estimator show a similar decrease in OMRP at small sample sizes, while bias and coverage probability remain unaffected. 
Because overlapping may increase computational burden in the stochastic optimization setting, we have examined computational efficiency of OMRP by jointly considering computation time and variance. 
Our results indicate that when warm starting is effective, efficiency can be observed across a wide range of overlapping degrees, and when warm starting is less effective, efficiency increases for only small degrees of overlap. 
We have empirically demonstrated computational efficiency by solving a newsvendor problem using insertion sort.
Finally, based on our study of computational efficiency, we have provided guidelines on selecting the degree of overlap in OMRP.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors thank Andrzej Ruszczy{\'{n}}ski and Artur {\'{S}}wietanowski for
access to their regularized decomposition code, which was used to solve CEP1 and DB1. 
The authors are also grateful to the associate editor and two anonymous referees for valuable suggestions that significantly improved both the content and the presentation of the paper. 
%This research has been supported in part by the National Science Foundation under Grant(s) DMS-0602173  EFRI-0835930 and CMMI-1345626 ??? 
An earlier version of this paper appeared in \citep{love2011overlapping}.

>>>>

\begin{figure}[thb!]
	\centering
	\subfloat[][Theoretical]{
	   \label{fig:efficiency_theory}
	   \includegraphics[width=.49\textwidth]{images/comp_efficiency_theory}}
	\subfloat[][Empirical for NVD]{
	   \label{fig:efficiency_nv}
	   \includegraphics[width=.51\textwidth]{images/nv_comp_efficiency}}
		\caption{
		Computational efficiency of the OMRP variance estimator relative to the (nonoverlapping) MRP variance estimator measured
		(a) theoretically and
		(b) empirically for NVD using insertion sort
%		(graphs are scaled so that nonoverlapping batches produces a computational efficiency of 1)
		}
\label{fig:efficiency}
\end{figure}

@@

We have applied the overlapping batch means method from simulation output analysis \citep{Meketon1984,Song1992,Welch1987} to assessing solution quality in stochastic programming, in particular, to Monte Carlo simulation-based estimators of optimality gaps \citep{Mak1999}. 
We have provided conditions under which (i) the resulting point estimator of the optimality gap and its associated variance estimator are consistent, (ii) OMRP variance estimators achieve asymptotically lower variances relative to the MRP variance estimator, and (iii) OMRP CIs are asymptotically valid. 
Empirical results indicate that asymptotic reductions in variance of the variance estimator show a similar decrease in OMRP at small sample sizes, while bias and coverage probability remain unaffected.

Overlapping may increase the computational burden in the stochastic optimization setting.
By jointly considering computation time and variance, we have examined the computational efficiency of OMRP, and we have empirically demonstrated  efficiency by solving a newsvendor problem using insertion sort.
Our analysis illustrates that when warm starting is effective, computational efficiency gains can be observed across a wide range of overlapping degrees.
Even when warm starting is less effective, efficiency slightly increases for low degrees of overlap. 
Maximizing efficiency and then minimizing total computation time suggests selecting the lowest degree of overlap across all the degrees with the highest efficiency for a given problem and solution method (with warm starting) combination.

%Finally, based on our study of computational efficiency, we have provided guidelines on selecting the degree of overlap in OMRP.

Overall, the results in this paper suggest that OMRP could be an efficient method to estimate optimality gaps for a class of stochastic programs when warm starting is available and/or when generating the observations is costly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors thank Andrzej Ruszczy{\'{n}}ski and Artur {\'{S}}wietanowski for
access to their regularized decomposition code, which was used to solve CEP1 and DB1. 
The authors are grateful to the associate editor and two anonymous referees for valuable suggestions that significantly improved both the content and the presentation of the paper. 
%This research has been supported in part by the National Science Foundation under Grants DMS-0602173 and CMMI-1345626.
%EFRI-0835930
An earlier version of this paper appeared in \citep{love2011overlapping}.
